---
title: "Memo - Exploring pre-1935 exposure data"
output:
   html_notebook:
    toc: true
    toc_float: true
---

## Purpose

The goal of this notebook is to explore initial exposure data for HOLC neighborhoods across the U.S. for wells with production dates on or before 12/31/1934 (approximately the time of HOLC grading).


```{r setup, include=FALSE}
# first set up the environment; we need the `tidyverse` and `sf` packages
source("../../code/0-setup/01-setup.R")
```

```{r data_input, include=FALSE}
# HOLC and wells data
holc_cities <- 
  st_read("../../data/raw/univ_richmond/mapping_inequality/shapefile/holc_ad_data.shp") %>%
  as_tibble() %>%
  mutate(city = ifelse(city == "Bronx" | city == "Brooklyn" | 
                         city == "Manhattan" | city == "Staten Island" |
                         city == "Queens", "New York City", city )) %>%
  mutate(city = as.factor(city),
         state = as.factor(state)) %>%
  group_by(city, state) %>%
  summarize()
holc_neighborhoods <- 
  st_read("../../data/raw/univ_richmond/mapping_inequality/shapefile/holc_ad_data.shp") %>%
  mutate(city = as.factor(city)) %>%
  st_transform(crs_projected) %>%
  st_buffer(dist = 0) %>%  # shortcut to address issue with dataset
  st_transform(crs_nad83) %>%
  mutate(neighborhood = as.character(neighborho)) %>%
  distinct(neighborhood, .keep_all = TRUE) %>%
  select(-neighborho)
wells_interim <- readRDS("../../data/interim/wells_interim.rds") %>%
  filter(spud_date <= as.Date("1934-12-31") | 
           completion_date <= as.Date("1934-12-31") |
           first_prod_date <= as.Date("1934-12-31") |
           last_prod_date  <= as.Date("1934-12-31"))

# exposure data
holc_cities_buffer_1km_wells_all <- 
  read_csv("../../data/interim/holc_cities_buffer_1km_wells_all_pre1935.csv")
holc_cities_buffer_3km_wells_all <- 
  read_csv("../../data/interim/holc_cities_buffer_3km_wells_all_pre1935.csv")

holc_neighborhoods_buffer_1km_wells_all <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_1km_wells_all_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))
holc_neighborhoods_buffer_3km_wells_all <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_3km_wells_all_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))

holc_neighborhoods_buffer_1km_wells_oil_gas <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_1km_wells_oil_gas_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))
holc_neighborhoods_buffer_3km_wells_oil_gas <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_3km_wells_oil_gas_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))

holc_neighborhoods_buffer_1km_wells_injection <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_1km_wells_injection_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))
holc_neighborhoods_buffer_3km_wells_injection <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_3km_wells_injection_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))

holc_neighborhoods_buffer_1km_wells_producing <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_1km_wells_producing_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))
holc_neighborhoods_buffer_3km_wells_producing <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_3km_wells_producing_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))

holc_neighborhoods_buffer_1km_wells_unknown <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_1km_wells_unknown_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))
holc_neighborhoods_buffer_3km_wells_unknown <- 
  read_csv("../../data/interim/holc_neighborhoods_buffer_3km_wells_unknown_pre1935.csv") %>%
  mutate(neighborhood = as.factor(neighborhood))
```

```{r data_prep, include=FALSE}
holc_neighborhoods_exp <- holc_neighborhoods %>%
  as_tibble() %>%
  mutate(neighhborhood = as.factor(neighborhood)) %>%
  left_join(holc_neighborhoods_buffer_1km_wells_all, by = "neighborhood") %>%
  mutate(buffer_1km_wells_all_n = replace_na(buffer_1km_wells_all_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_3km_wells_all, by = "neighborhood") %>%
  mutate(buffer_3km_wells_all_n = replace_na(buffer_3km_wells_all_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_1km_wells_oil_gas, by = "neighborhood") %>%
  mutate(buffer_1km_wells_oil_gas_n = replace_na(buffer_1km_wells_oil_gas_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_3km_wells_oil_gas, by = "neighborhood") %>%
  mutate(buffer_3km_wells_oil_gas_n = replace_na(buffer_3km_wells_oil_gas_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_1km_wells_injection, by = "neighborhood") %>%
  mutate(buffer_1km_wells_injection_n = replace_na(buffer_1km_wells_injection_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_3km_wells_injection, by = "neighborhood") %>%
  mutate(buffer_3km_wells_injection_n = replace_na(buffer_3km_wells_injection_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_1km_wells_producing, by = "neighborhood") %>%
  mutate(buffer_1km_wells_producing_n = replace_na(buffer_1km_wells_producing_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_3km_wells_producing, by = "neighborhood") %>%
  mutate(buffer_3km_wells_producing_n = replace_na(buffer_3km_wells_producing_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_1km_wells_unknown, by = "neighborhood") %>%
  mutate(buffer_1km_wells_unknown_n = replace_na(buffer_1km_wells_unknown_n, 0)) %>%
  left_join(holc_neighborhoods_buffer_3km_wells_unknown, by = "neighborhood") %>%
  mutate(buffer_3km_wells_unknown_n = replace_na(buffer_3km_wells_unknown_n, 0))
```


## Cities

There were `r nrow(holc_cities)` cities in the HOLC dataset. Of these, `r holc_cities_buffer_1km_wells_all %>% filter(buffer_1km_wells_all_n > 0) %>% nrow()` had wells that were drilled or in operation before 1935 within 1 km a HOLC-graded neighborhood. Of these, `r holc_cities_buffer_1km_wells_all %>% filter(buffer_1km_wells_all_n >= 10) %>% nrow()` had at least 10 wells.

#### Table. Number of wells within 1 km of graded neighborhoods

```{r echo=FALSE}
holc_cities_buffer_1km_wells_all %>%
  filter(buffer_1km_wells_all_n >= 10) %>%
  arrange(desc(buffer_1km_wells_all_n))
```


## Wells

A total of `r nrow(wells_pre1935_sf)` wells had a spud date, completion date, and/or production start or end date before 1935. Many of the wells with production dates pre-1935 were of unknown type.

```{r echo=FALSE, message=FALSE, warning=FALSE}
wells_interim %>%
  mutate(production_type = as.factor(production_type)) %>%
  group_by(production_type) %>%
  summarize(n = n()) %>%
  ggplot() +
  geom_bar(aes(reorder(production_type, n), n), stat = "identity") +
  theme_classic() +
  coord_flip()
```


## Exposure to all wells

Unit of observation is the neighborhood 

### All wells within 1 km of neighborhood centroid

#### Plot. Histogram of n all wells within 1 km 

Restricted to neighborhoods with > 0 wells. Note right skew in the data; most neighborhoods were unexposed.

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_all_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_1km_wells_all_n), bins = 100) +
  labs(x = "All wells within 1 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n all wells within 1 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_all_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "All wells within 1 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n all wells within 1 km, by city

Most pre-1935 HOLC neighborhoods with wells were in CA, OH, and OK.

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_all_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_all_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "All wells within 1 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```


### All wells within 3 km of neighborhood centroid

#### Plot. Histogram of n all wells within 3 km 

Restricted to neighborhood with > 0 wells

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_3km_wells_all_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_3km_wells_all_n), bins = 100) +
  labs(x = "All wells within 3 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n all wells within 3 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_3km_wells_all_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "All wells within 3 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n all wells within 3 km, by city

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_3km_wells_all_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_3km_wells_all_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "All wells within 3 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```


## Exposure to oil and gas wells

Unit of observation is the neighborhood 

### Oil and gas wells within 1 km of neighborhood centroid

#### Plot. Histogram of n oil and gas wells within 1 km 

Note: Right skew in the data; most neighborhoods were unexposed.

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_oil_gas_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_1km_wells_oil_gas_n), bins = 100) +
  labs(x = "Oil and gas wells within 1 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n oil and gas wells within 1 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_oil_gas_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "All wells within 1 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n oil and gas wells within 1 km, by city

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_all_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_oil_gas_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "All wells within 1 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```


## Exposure to injection wells

Only 1 neighborhood had an injection well within 1 km and 29 had at least one well within 3 km. I'll exclude injection wells for the pre-1935 analyses.


## Exposure to producing wells 

Producing wells were defined as wells with > 0 BOE of oil and/or gas production, as reported in the Enverus dataset.

### Producing wells within 1 km of neighborhood centroid

#### Plot. Histogram of n producing wells within 1 km 

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_producing_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_1km_wells_producing_n), bins = 100) +
  labs(x = "Producing wells within 1 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n producing wells within 1 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_producing_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Producing wells within 1 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n producing wells within 1 km, by city

Note: All data for California come from Los Angeles, I think that will make an interesting case study!

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_producing_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_producing_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Producing wells within 1 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```


### Producing wells within 3 km of neighborhood centroid

#### Plot. Histogram of n producing wells within 3 km 

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_3km_wells_producing_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_3km_wells_producing_n), bins = 100) +
  labs(x = "Producing wells within 3 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n producing wells within 3 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_3km_wells_producing_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Producing wells within 3 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n producing wells within 3 km, by city

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_3km_wells_producing_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_3km_wells_producing_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Producing wells within 3 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```


## Exposure to 'unknown' wells

Some `r wells_pre1935_sf %>% filter(production_type == "UNKNOWN") %>% nrow()` of the pre-1935 wells were labeled as unknown type. Given the age of these records, I'm not too surprised to see this.

### Unknown wells within 1 km of neighborhood centroid

#### Plot. Histogram of n unknown wells within 1 km 

Note: Right skew in the data; most neighborhoods were unexposed.

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_unknown_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_1km_wells_unknown_n), bins = 100) +
  labs(x = "Unknown wells within 1 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n unknown wells within 1 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_unknown_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Unknown wells within 1 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n oil and gas wells within 1 km, by city

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_1km_wells_all_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_1km_wells_unknown_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Unknown wells within 1 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```


### Unknown wells within 3 km of neighborhood centroid

#### Plot. Histogram of n unknown wells within 3 km 

Note: Right skew in the data; most neighborhoods were unexposed.

```{r echo=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_3km_wells_unknown_n > 0) %>%
  ggplot() +
  geom_histogram(aes(buffer_3km_wells_unknown_n), bins = 100) +
  labs(x = "Unknown wells within 3 km of neighborhood centroid (n)",
       y = "Count") +
  theme_classic()
```

#### Plot. Histogram of n unknown wells within 3 km, all cities

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_3km_wells_unknown_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Unknown wells within 3 km of neighborhood centroid (n)") +
  theme_classic()
```

#### Plot. Histogram of n oil and gas wells within 3 km, by city

Most of these wells seem to be in Oklahoma City and Tulsa, where we see a disparity in exposure.

```{r echo=FALSE, message=FALSE, warning=FALSE}
holc_neighborhoods_exp %>%
  filter(buffer_3km_wells_all_n > 0) %>%
  mutate(holc_grade = as.factor(holc_grade)) %>%
  filter(holc_grade %in% c("A", "B", "C", "D")) %>%
  ggplot() +
  geom_bar(aes(holc_grade, buffer_3km_wells_unknown_n), stat = "identity") +
  labs(x = "HOLC grade",
       y = "Unknown wells within 3 km of neighborhood centroid (n)") +
  facet_wrap(. ~ city) +
  theme_classic()
```